Motion planning for assessing accessibility of protein tunnels

### 2017-09-07 paper.tex
added images, new intro image

\begin{figure}[b]
\centering
{\footnotesize
\renewcommand{\arraystretch}{0.1}
\renewcommand{\tabcolsep}{0pt}
\begin{tabular}{ccc}
\includegraphics[width=0.15\textwidth]{fig/motiv1} &
\includegraphics[width=0.17\textwidth]{fig/motiv2lab} &
\includegraphics[width=0.16\textwidth]{fig/motiv3}  \\
%\hbox{
%\vbox{
%\hbox{\includegraphics[width=0.25\textwidth]{fig/ta-1} }
%\hbox{\includegraphics[width=0.25\textwidth]{fig/ta-433}}
%} 
%}
Protein 1CQW & Active site & Detected tunnels \\ %& Example of a  \\
             &            & (orange)          %& trajectory
\end{tabular}
}
\caption{\label{fig::motiv}
    Tunnels in haloalkane dehalogenase with a possible trajectory of 1-chlorpropan ligand.
}
\end{figure}

The idea of sampling-based motion planning is to randomly sample the configuration space of the robot and connect the collision-free samples
into a graph structure (roadmap).
A path in the roadmap then corresponds to a motion in the workspace.
The random samples are classified as free or non-free using black-box collision detection which allows the sampling-based methods to find paths for robots of various shapes. 
The sampling-based planners is that they can cope with robots with many degrees of freedom (DOF).
The two most used sampling-based techniques are Probabilistic Roadmaps (PRM)~\cite{kavrakiForPP} and Rapidly Exploring Random Trees (RRT)~\cite{lavalleRRT}.
Besides many applications in robotics, sampling-based planners have been used e.g. in bio-chemistry 
%Sampling-based motion planners has been intensively studied in robotics, but they have been applied also in other research fields.
%In bio-chemistry, the planners have been used to study
to study loop motions~\cite{cortes2004geometric},
%protein folding~\cite{amato2002using,raveh2009rapid,novinskaya2015improving,songPFintro},
or for protein folding~\cite{raveh2009rapid,novinskaya2015improving}.
%or for tunnel detection~\cite{vonasek2017tunnel}.

One of the problems in protein engineering is the tunnel detection, where the task is to find a tunnel (pathway) 
leading to a specific place (called active site) inside the protein (MAME CITACI?)
%or protein folding combined with ligand diffusion~\cite{cortes2010simulating}.
%rrt for molecular: \cite{al2012motion}
%survey \cite{gipson2012computational}
%Sampling-based planners have been used in various applications in robotics~\cite{elbanhawi2014sampling} and also  %,latombe1999motion} and also
%Solutions for tunnel detection using sampling-based methods however have not been discussed yet.
%Understanding of interactions between proteins and other small molecules is crucial in many research fields, including drug design and protein engineering. 
The knowledge about the existence of the tunnels and their properties (e.g., their bottleneck and chemical properties around the tunnel),
is important to understand interactions between proteins and other molecules (ligands)~\cite{Koudelakova2013}.
To get the ligand to the active site, a transportation tunnel connecting the protein outer environment with the active site has to exists.
Such a  tunnel has to be wide enough and the physico-chemical properties of the surrounding amino acids have to be compatible with the ligand (Fig.~\ref{fig::motiv}).
%These interactions are highly influenced by the ability to transport the small molecule, called ligand, to the protein active site.
%The active site can be considered as a deeply buried inner cavity with the ability to interact with the incoming ligand.
The tunnels are usually detected using Voronoi diagrams~\cite{yaffe2008,caver3} assuming 
a spherical ligand (probe) and they are represented as a sequence of spheres.
However, as the ligands typically non-spherical and flexible,  it is difficult to estimate their traversability through the tunnels computed for only one sphere.

>>>>

\begin{figure}[b]
\centering
{\footnotesize
\renewcommand{\arraystretch}{0.1}
\renewcommand{\tabcolsep}{0pt}
\begin{tabular}{ccc}
\includegraphics[width=0.15\textwidth]{fig/motiv1} &
\includegraphics[width=0.17\textwidth]{fig/motiv2lab} &
\includegraphics[width=0.16\textwidth]{fig/motiv3}  \\
Protein 1CQW & Active site & Detected tunnels \\ %& Example of a  \\
             &            & (orange)         \\  %& trajectory
\multicolumn{3}{c}{%
\includegraphics[width=0.16\textwidth]{fig/renderDCP} 
\includegraphics[width=0.16\textwidth]{fig/render37t}} \\ 
%\hbox{
%\vbox{
%\hbox{\includegraphics[width=0.25\textwidth]{fig/ta-1} }
%\hbox{\includegraphics[width=0.25\textwidth]{fig/ta-433}}
%} 
%}
\end{tabular}
}
\caption{\label{fig::motiv}
    Tunnels in haloalkane dehalogenase with a possible trajectory of 1-chlorpropan ligand.
}
\end{figure}

The idea of sampling-based motion planning is to randomly sample the configuration space of the robot and connect the collision-free samples
into a graph structure (roadmap).
A path in the roadmap then corresponds to a motion in the workspace.
%The random samples are classified as free or non-free using black-box collision detection which allows the sampling-based methods to find paths for robots of various shapes. 
%The sampling-based planners can cope with robots with many degrees of freedom (DOF).
The two most used sampling-based techniques are Probabilistic Roadmaps (PRM)~\cite{kavrakiForPP} and Rapidly Exploring Random Trees (RRT)~\cite{lavalleRRT}.
The sampling-based planners can cope with many-DOF (Degrees of Freedom) robots of arbitrary shape and therefore they have been used for motion planning of varous robotic systems.
Besides many applications in robotics, sampling-based planners have been used e.g. in bio-chemistry 
%Sampling-based motion planners has been intensively studied in robotics, but they have been applied also in other research fields.
%In bio-chemistry, the planners have been used to study
to study loop motions~\cite{cortes2004geometric},
%protein folding~\cite{amato2002using,raveh2009rapid,novinskaya2015improving,songPFintro},
or for protein folding~\cite{raveh2009rapid,novinskaya2015improving}.
%or for tunnel detection~\cite{vonasek2017tunnel}.

One of the problems in protein engineering is the tunnel detection, where the task is to find a tunnel (pathway) 
leading to a specific place (called active site) inside the protein~\cite{gora2013gates}.
%or protein folding combined with ligand diffusion~\cite{cortes2010simulating}.
%rrt for molecular: \cite{al2012motion}
%survey \cite{gipson2012computational}
%Sampling-based planners have been used in various applications in robotics~\cite{elbanhawi2014sampling} and also  %,latombe1999motion} and also
%Solutions for tunnel detection using sampling-based methods however have not been discussed yet.
%Understanding of interactions between proteins and other small molecules is crucial in many research fields, including drug design and protein engineering. 
The knowledge about the existence of the tunnels and their properties (e.g., their bottleneck and chemical properties around the tunnel),
is important to understand interactions between proteins and other molecules (ligands)~\cite{Koudelakova2013}.
To get the ligand to the active site, a transportation tunnel connecting the protein outer environment with the active site has to exists.
Such a  tunnel has to be wide enough and the physico-chemical properties of the surrounding amino acids have to be compatible with the ligand (Fig.~\ref{fig::motiv}).
%These interactions are highly influenced by the ability to transport the small molecule, called ligand, to the protein active site.
%The active site can be considered as a deeply buried inner cavity with the ability to interact with the incoming ligand.
The tunnels are usually detected using Voronoi diagrams~\cite{yaffe2008,caver3} assuming 
a spherical ligand (probe) and they are represented as a sequence of spheres.
However, as the ligands typically non-spherical and flexible,  it is difficult to estimate their traversability through the tunnels computed for only one sphere.
